---
title: "EmoRy Coding Club"
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# When & Where

**Date:** July 11, 2019, 04:00 - 6:00p  
**Location:** Rm 2052, O. Rollins


<br>


# Before you arrive:

Please complete the following install & setup.

To participate in this workshop, you will need to make sure you:  

* Install [R](https://www.r-project.org/) from CRAN. 
  + If you're asked to choose a CRAN mirror, just choose whichever is closest to your location.
* Install [RStudio Desktop](https://www.rstudio.com/products/rstudio/download/)
* Install the `tidyverse` package and then load it by copying and pasting the following code into the *Console* of RStudio and then pressing *Enter* to execute it.
```{r, eval= FALSE}
install.packages("tidyverse")
install.packages("ggplot2")
library(tidyverse)
library(ggplot2)
```


<br>


# Pre-work questions
If you you have no previous experience with R or coding, then before arriving please complete the pre-work below to get familiar with a few basic things about the R Studio IDE and coding in R. The exercise below should take less than 5 minutes to complete.

Once you have installed R and R Studio, copy and paste the code below into the R Studio *Console* and press *Enter* to run it.

```{r, eval = FALSE}

# Caculate something simple
2 + 2

# Calculate something using built-in contants
2*pi

```

<br>

Use the *assignment operator*, `<-`, to create new objects in R

```{r, eval = FALSE}

a <- 2 + 2
b <- a*pi
c <- "Hello World"
```

And then print the objects in the console:

```{r, eval = FALSE}
a 
b
c
```


# Learning Objectives

Our first EmoRy Coding Club will be primarily targeted towards beginners, but we hope this will also useful for anyone who may already be familiar with base R but would like to get better practice with the R `tidyverse`.

1) Introducing beginners to the RStudio IDE
2) Doing some introductory plotting with `ggplot`

We will use the [R for Data Science](https://r4ds.had.co.nz/index.html) free online text to guide us through learning R.


<br>

# Materials

We'll be diving into [Ch 3, Data visualisation](https://r4ds.had.co.nz/data-visualisation.html) and working through Section 3.3.1.


Once we've completed the R for Data Science exercises, we'll repeat the plotting exercise using open-access data from AirBnb.

This is a really big data set (it has ~50k rows !!), so it will take a moment to run. To import the AirBnb data from the URL, copy and paste the lines of code below:

```{r, eval = FALSE}
require(readr)

# function for reading in zipped file (106 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2019-06-02/data/listings.csv.gz"

url_read <- function(file_url) {
  con <- gzcon(url(file_url))
  txt <- readLines(con)
  return(read.csv(textConnection(txt)))
}

df <- url_read(url) # read in data

# smaller csv file (16 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2019-06-02/visualisations/listings.csv"
df <- read.csv(url)

```


```{r, echo =FALSE, }

# devtools::install_github('yihui/xfun')

# a single RmD file
# xfun::embed_file('example.Rmd')

# embed a single csv file
 # r xfun::embed_file('data.csv')` 

```

