---
params:
  lesson: "Lesson 3"
  title: "Using `filter`, `arrange`, and `R` project (`Rproj`) files"      
  bookchapter_name: "Section 4"    
  bookchapter_section: "https://r4ds.had.co.nz/workflow-basics.html"
  functions: "`p_load`, `==`, `|`, `%in%`, `filter`, `near`"
  packages: "`ggplot2`,`pacman`, `dplyr`, `nycflights13`"      
  # end inputs --------------------------------------------------------------- 
header-includes: \usepackage{float}
always_allow_html: yes
output:
  html_document:
    code_folding: show
---
  
```{r, setup, echo = FALSE, cache = FALSE, include = FALSE}
options(width=100)
knitr::opts_chunk$set(
  eval = FALSE, # run all code
  echo = TRUE, # show code chunks in output 
  tidy = TRUE, # make output as tidy
  message = FALSE,  # mask all messages
  warning = FALSE, # mask all warnings 
  comment = "",
  tidy.opts=list(width.cutoff=100), # set width of code chunks in output
  size="small" # set code chunk size
  )
```
\

<!-- install packages -->
```{r, load packages, eval=T, include=T, cache=F, message=F, warning=F, results='hide',echo=F}
packages <- c("ggplot2","ggthemes","dplyr","tidyverse","zoo","RColorBrewer","viridis","plyr", "nycflights13")   
if (require(packages)) {
    install.packages(packages,dependencies = T)
    require(packages)
    # load tvthemes
    devtools::install_github("Ryo-N7/tvthemes")
}
lapply(packages,library,character.only=T)  
```

<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- start body -->

# `r paste0(params$lesson,": ",params$title)`    
\  

Functions for `r params$lesson`  
`r params$functions`    
\    

Packages for `r params$lesson`          
`r params$packages`        
\    

# Agenda

[Data visualisation in `R` for Data Science, `r params$bookchapter_name`](`r params$bookchapter_section`).      
\  

<!--  end yaml template------------------------------------------------------- -->
* Learn how to use R project files  
* Use `pacman` for easy package loading  
* Solidify some fundamental `R` operators  
* Subset data using `dplyr::filter()`

# Do First    

Recreate the below plot using the smaller NYC Airbnb dataset.  

```{r}
# smaller csv file (16 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2021-04-07/data/listings.csv.gz"
nyc <-  read_csv(url)
nyc <- nyc[nyc$id < 20000,] # get smaller subet of data
```

```{r, echo=F, eval=T, out.width="100%"}
require(ggthemes)
require(ggplot2)
require(readr)

# smaller csv file (16 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2021-04-07/data/listings.csv.gz"
nyc <-  read_csv(url)
nyc <- nyc[nyc$id < 10000,] # get smaller subet of data

ggplot(data = nyc, mapping = aes(x = minimum_nights, y = price)) +
  geom_point(aes(color=neighbourhood_group_cleansed),show.legend = F) +
  facet_wrap(~ neighbourhood_group_cleansed, nrow=length(nyc$neighbourhood_group_cleansed)) +
  theme_classic()

```
\  

# Load packages with `pacman`  
```{r}
install.packages("pacman")
require(pacman)
```  

Type in the packages you want to install into the `p_load` function.  
_Note: the below two arguments are the same_
```{r}
p_load(dplyr,ggplot2,nycflights13) # your packages  
```

```{r}
pacman::p_load(dplyr,ggplot2,nycflights13) # this is the same as above, but allows you to use a specific function from a package without loading the entire package: 'package::function(argument1,argument2,argument3)'  
```
\  

# Introducing `R` project (`Rproj`) files.  
You can set individual working directories and contain your R documents related to a specific project using an `Rproj` file.  Two options:

**Option 1:** Create an RProject new directory:
1. _File > New Project_
1. _Create New Project_  
1. Choose a name for your RProj folder. Fill out _Directory name:_ (make it machine-friendly, i.e. no spaces).
1. Choose a place for the RProj to live. _Browse_
1. Select _Open in new session_    
\    

**Option 2:** If you already have a folder *just* for Emory Coding Club:
1. _File > New Project_
2. _Existing directory > Browse_  
3. Select _Open in new session_    

\    

Open a new `R` file within this project and save it as 'Lesson3.R'.  
\    

Now repeat the above steps so that you have two different `Rproj` files. Name this second project something different. 
\    

Let's now see why `Rproj` files are a useful workflow method by checking the working directory for each project.  
```{r}
getwd()
```
\  

# Basics in `R`  
Learning and applying the fundamentals of `R` so you can build more complex analytical tools. Use the below summary sheet of some common, basic `R` operations to solve the practice questions in the following section.        
\  

## Summary sheet of `R` fundamentals      
```{r}
# numeric calculations 
1 / 200 * 30
#> [1] 0.15
(59 + 73 + 2) / 3
#> [1] 44.7
sin(pi / 2)
#> [1] 1

# -------------------------------------------

# assigning objects 
require(tidyverse)  
x <- 4 * 3 # conventional method  
3 * 4 -> x # occasional tidyverse method to be aware of  
x %>% sqrt -> x # used with a pipe ( %>% )

# -------------------------------------------

# missing values (NAs)  
NA > 5
10 == NA
NA + 10
NA / 2

NA == NA # get NA
x <- NA
y <- NA
x == y # x and y are both unknowns, so the above returns NA

# check if NAs exist  
is.na(x)

# -------------------------------------------

# conditionals 
5 != 5 # is not equal to 
5 < 10
5 > 10
5 >= 10
5 <= 10

```
\  

# [Chapter 5: Transformations](https://r4ds.had.co.nz/transform.html)    
Load the datasets    
```{r}
library(nycflights13)
flights # print the flights dataset

```
\  

# Data classes   

`int` stands for integers.  

`dbl` stands for doubles, or real numbers.  

`chr` stands for character vectors, or strings.  

`dttm` stands for date-times (a date + a time).  

`lgl` stands for logical, vectors that contain only TRUE or FALSE.  

`fctr` stands for factors, which R uses to represent categorical variables with fixed possible values.  

`date` stands for dates.  
\  

# Using `dplyr`  

## `filter()`  
Pick observations by their values  
```{r}
jan1 <- filter(flights, month == 1, day == 1)
(jan1 <- filter(flights, month == 1, day == 1)) # wrap in parentheses to both assign and execute  

filter(flights, month = 1)  
```
\  

## Logical operators    

The following code finds all flights that departed in November or December.  
```{r}
glimpse(flights)

# using `|`
filter(flights, month == 11 | month == 12)

# using `%in%`  
nov_dec <- filter(flights, month %in% c(11, 12)) 

# a '!' specifies a range WITHOUT those conditions, i.e. equal to not including those values 
filter(flights, !(arr_delay > 120 | dep_delay > 120)) 
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
\    

Retaining NA values    
```{r}
df <- tibble(x = c(1, NA, 3)) # create tibble dataframe  
filter(df, x > 1) # omit NA by default
filter(df, is.na(x) | x > 1) # retain NA

```
\  

# Exercise 5.2.4 
[Try the exercises from 5.2.4](https://r4ds.had.co.nz/transform.html#exercises-7).  
\  

# Changing order of data rows with `arrange()`  

Change rows to ascending order  
```{r}
arrange(flights, year, month, day)
```
\  

`desc()` = descending order  
```{r}
arrange(flights, desc(dep_time))
```
\  

This puts missing values at the end of the dataframe  
```{r}
df <- tibble(x = c(5, 2, NA)) # create mock data 
arrange(df, x)
arrange(df, desc(x))
```
\  

# Exercise 5.3.1 
Try the ['5.3.1 Exercises' using `arrange()`](https://r4ds.had.co.nz/transform.html#exercises-8)  
\  

# Applying the Airbnb data  
Use the new examples on the Airbnb dataset.        
```{r}
# large Airbnb dataset (106 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2021-04-07/data/listings.csv.gz"

nyc_full <- read_csv(url) # reads in data
nyc_full %>% glimpse
```

<!-- end body -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
