---
title: "_lesson4_practice"
author: "Desirée De Leon"
date: "9/29/2019"
output: html_document
---

```{r}
start <- Sys.time()

#load packages

library(nycflights13)
library(tidyverse)

```

#5.2.3 Missing Values

```{r}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
```
```{r}
filter(df, is.na(x) | x > 1)
```



#5.3 Arrange rows with `arrange()`

#5.3.1 Exercises

1. How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()).

```{r}

df <- tibble(x = c(5, 2, NA), y= c(1, 2, 3))
arrange(df, desc(y), x)



a <- arrange(df, desc(is.na(x))) #NA, biggest to smallest b/c TRUE = 1, and FALSE=0

#but what about smallest to biggest with NA leading?
b <- arrange(df, desc(is.na(x)))




```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.


```{r}
# most delayed

flights # use dep_delay as var

arrange(flights, desc(dep_delay)) #delay by departure (not arrival)

#left the earliest

arrange(flights, dep_delay) #neg numb means early, which I knew because of ?flights

```

3. Sort flights to find the fastest flights.

```{r}
#fastest as in distance/ time?
arrange(flights, distance/air_time)
```


4. Which flights travelled the longest? Which travelled the shortest?

```{r}
# the longest?
arrange(flights, desc(air_time)) # EWR (newark) to HNL @ 695 minute air_time

# the shortest
arrange(flights, air_time) # Newark to Connecticut EWR BDL @ 20 minutes

```

# 5.4 Select columns with select()
`select` is like subsetting in base r

# 5.4.1 Exercises

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
# start w/ glimpse again
#

#1) by straight up listing them
select(flights, dep_time, dep_delay, arr_time, arr_delay) 

#2)  using "starts_with"
select(flights, starts_with("dep"), starts_with("arr"))

#3) using : and then excluding the stuff you don't want in between

select(flights, dep_time:arr_delay, -(starts_with("sched")))

#4) using contains

select(flights, contains("dep"), starts_with("arr"), -contains("sched"))

```

2. What happens if you include the name of a variable multiple times in a select() call?
 - nothing??
```{r}

select(flights, dep_time, dep_time, arr_time)
```

3. What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?
`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`

The `one_of` function can be used so that you can select any variables that are within `one_of`. 

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay", "hi") # includes bogus column name

select(flights, vars) #does not work 
select(flights, one_of(vars)) # but this does

```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))

```

Yes, because of the caps in "TIME". Run `?contains()` to see why... `contains()` has a default setting of `ignore.case = TRUE`. Can change to `= FALSE`.

```{r}
select <- Sys.time()
select-start
```

# 5.5 Add new variables with mutate()

## 5.5.2

1. Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}

transmute(flights, 
       dep_time, 
       dep_time_min = dep_time %/% 100 * 60 + dep_time %% 100, 
       sched_dep_time, 
       sched_dep_time_min = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100, 
         )



# but also midnight (2400) would become 1400 minutes instead of 0, so need to convert that too

transmute(flights, 
       dep_time, 
       dep_time_min = dep_time %/% 100 * 60 + dep_time %% 100, 
       sched_dep_time, 
       sched_dep_time_min = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100, 
         ) %% 1400

```

2. Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}

transmute(flights, air_time,
          arr_time_min = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
          dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440, 
          arr_time_min - dep_time_min)

# you expect to see the same values, but you don't. Need to use the "easy to compute" versions of arr_time and dep-time
# also, time zones.  but not multiples of 60. Also, arrival time could include time to taxi to/fro gate but is not airtime
# check documentation, data source. See we don't have variables in this data set like "taxiin" taxioff" wheelsin, wheelsoff"

# Solution:
# if it was only a time zone issue, would convert everything to a common time zone (GMT)
```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

- you expect scheduled time + delay = dep time

```{r}

transmute(flights, dep_time, sched_dep_time, dep_delay)
# but needs to be in continuous var form
```


```{r}
# save in new tibble to make easier. make new column variables.
dep <- mutate(flights,
                 dep_delay,
                 sched_dep_time_min = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440,
                 dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
                 diff = sched_dep_time_min + dep_delay - dep_time_min)

dep
```

```{r}
# are there any obs where diff != 0?

filter(select(dep, contains("dep"), diff) , diff != 0)

# yes, and all the discrepancies are the non-0 vals are 1440 (which would be for scheduled flights before midnight that got delayed to be after midnight) but let's ggplot to be sure.
```

ggplot to see 

```{r}

ggplot(data = filter(dep, diff !=0), mapping = aes(y = sched_dep_time_min, x = diff)
) + 
  geom_point()

```

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for `min_rank()`.

- min rank gives tied values the same rank, but then skips how ever many ties there were to get the next rank
- dense rank, picks up where the ties left off


```{r}
# 10 most delayed, but use a rank function instead of arrange()

# first, rank stuff
ranked_delays <- mutate(flights, dep_delay,
       rank_delay = min_rank(desc(dep_delay)))

# then, filter based on that

filter(ranked_delays, rank_delay <= 10) # no ties

```



5. What does 1:3 + 1:10 return? why?

```{r}
# learned about : with select() to grab columns between col:col and also for num_range("x", 1:3)
1:3 + 1:10
```
Gives warning and then output.
It's the same as saying:

1+1, 2+2, 3+3...1+4...The first range of numbers will recycle until it reaches the end of the longer object (i.e. 1:10)

6. What trigonometric functions does R provide?

This question seems so random...

 https://www.rdocumentation.org/packages/base/versions/3.6.1/topics/Trig

```{r}

#Hmm.....

sin(pi*1)
sinpi(1)


```




