---
title: EmoRy Coding Club master document for club sessions  
author: |
 | Matthew Malishev*, Desiree De Leon & Hasse Walum     
 |  
 | _Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
output:
  html_document:
    highlight: tango
    code_folding: show
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
params: 
  # dir: "/Users/malishev/Documents/Emory/admin/ecc/"
  date: !r Sys.Date()
  session: !r sessionInfo()  
  version: !r getRversion()
  email: "matthew.malishev@emory.edu"
  doi: https://github.com/darwinanddavis/EmoRyCodingClub
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
#knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

#setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\newpage   

Date: `r params$date`  
R version: `r params$version`  
This document can be found at `r params$doi`  

<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- start body -->

\newpage  

## Example of first session   

#### Pre session  
1. assign 'hello world' to variable    
2. print to console  
3. run help page for 'print'  

#### First session  
```{r,eval=F}
install.packages()  
head()  
tail()
str()

plot()

df$Col1
df[,"Col1"]

```

#### TOC    
What the session covers  
Dataset  
Showcase of what you'll learn by the end  

```{r,eval=F}
df <- read.csv("https://query.data.world/s/beephekfuniqs7n3ym2uitmbvoccgs", header=TRUE, stringsAsFactors=FALSE)

str(df)
unique(df$completed_on)[1]
plot(df$end_location_lat,df$end_location_long,pch=20,main="R plot",las=1,xlim=c(30,35),ylim=c(-112,98))

```


#### Importing Data

AirBnb Data

```{r, eval = FALSE}
require(readr)

# function for reading in zipped file (106 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2019-06-02/data/listings.csv.gz"

url_read <- function(file_url) {
  con <- gzcon(url(file_url))
  txt <- readLines(con)
  return(read.csv(textConnection(txt)))
}

df <- url_read(url) # read in data

# smaller csv file (16 cols)
url <- "http://data.insideairbnb.com/united-states/ny/new-york-city/2019-06-02/visualisations/listings.csv"
df <- read.csv(url)


```


**Before the first session, let's see if we can:**

1) Read this in so that it automatically prints out as a tibble (alternatively, use `tibble(df)` to print)
2) Reduce/ subset the number of rows to something closer to ~10k? to decrease time to load/ plot, etc. 



<!-- end body -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
<!-- ____________________________________________________________________________ -->
